---
title: "Network Assessment and Spatiotemporal Travel Patterns of Metro System in London"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "Network Analysis"
output: 
  html_document:
      toc: true
      toc_float: true
      collapsed: true
      number_sections: false
      toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Data Process

Transform and clean the MTT(Monday to Thursday) origin and destination data before analysis

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(tmap)
library(sf)
library(sp)
library(ggplot2)
library(shp2graph)
library(rgdal)
tmap_mode("view")
# Read in 
  ## For actual demand
OD_cleaned <- read_csv("Data/OD_cleaned.csv")
stations <- read_csv("Data/London stations2.csv")
London_Boundary <- st_read("Data/London boundary/ESRI/London_Borough_Excluding_MHW.shp")


  ## For estimated demand
EstiFlow <- readOGR("Data/London Train Lines.KML")

    ### Calculation with iGraph object
EstiFlow_IG1 <- readshpnw(EstiFlow,ELComputed =TRUE)
EstiFlow_IG2 <- nel2igraph(EstiFlow_IG1[[2]],EstiFlow_IG1[[3]],weight = EstiFlow_IG1[[4]])
PageRank <- page.rank(EstiFlow_IG2,weights = EstiFlow_IG1[[4]])

    ### Change back to dataframe
EstiFlow <- as_data_frame(EstiFlow_IG2,what ="vertices")
EstiFlow <- cbind(EstiFlow,PageRank$vector)
colnames(EstiFlow) <- c("Longitude","Latitude","PageRank")

    ### Convert to sp object
EstiCap <- SpatialPoints(EstiFlow[,c("Longitude","Latitude")],proj4string = CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))
EstiCap <- SpatialPointsDataFrame(EstiCap,EstiFlow[3])



# Caluclate the demand of each station
Alighter <- aggregate(OD_cleaned["Total"],by=OD_cleaned["Destination Station Name"], FUN=sum)
Boarder <- aggregate(OD_cleaned["Total"],by=OD_cleaned["Origin Station Name"], FUN=sum)
ActuFlow <- merge(Alighter,Boarder,by.x = "Destination Station Name",by.y ="Origin Station Name")
ActuFlow[4] <- ActuFlow[2]+ActuFlow[3]
ActuFlow <- ActuFlow[c(1,4)]




# Add coordinates
colnames(ActuFlow) <- c("Station Name","Total Flow")
ActuFlow <- merge(ActuFlow,stations,by.x = "Station Name",by.y = "Station")
```


## Create Spatial Object

```{r}
# Create spatialpointsdataframe
ActuCap <- SpatialPoints(ActuFlow[,c("Longitude","Latitude")],proj4string = CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))

ActuCap <- SpatialPointsDataFrame(ActuCap,ActuFlow[,1:2])
```

## Actual Demand
### Distribution of Actual Demand

```{r}
# Distribution
ActuCapDis <- ggplot(ActuFlow)+
  geom_histogram(aes(x = ActuFlow$`Total Flow`),
                 fill = brewer.pal(7,"Blues")[7],
                 col =brewer.pal(7,"Greys")[2],
                 alpha = 0.8,
                 bins = 60)+
  scale_x_continuous(breaks = seq(min(ActuFlow$`Total Flow`),
                                  max(ActuFlow$`Total Flow`),
                                  (max(ActuFlow$`Total Flow`)-min(ActuFlow$`Total Flow`))/10))+
  scale_y_continuous()+
  labs(title="Distribution of Actual Demand(capita)",
       x="Actual Demand(capita)",
       y="Frequency")




ActuCapDis
```


### Transformed Distribution of Actual Demand

```{r}
# Transformed Distribution
ActuCapDis_Tran <- ggplot(ActuFlow)+
  geom_histogram(aes(x = log10(ActuFlow$`Total Flow`)),
                 fill = brewer.pal(7,"Blues")[7],
                 col =brewer.pal(7,"Greys")[2],
                 alpha = 0.8,
                 bins = 60)+
  scale_x_continuous(breaks = seq(round(min(log10(ActuFlow$`Total Flow`)),2),
                                  round(max(log10(ActuFlow$`Total Flow`)),2),
                                  round((max(log10(ActuFlow$`Total Flow`))-min(log10(ActuFlow$`Total Flow`)))/10,1)))+
  scale_y_continuous()+
  labs(title="Distribution of log(Actual Demand(capita))",
       x="log10(Actual Demand(capita))",
       y="Frequency")




ActuCapDis_Tran
```


### Actual Demand on Map

```{r}
# Base map
London_Base <- tm_shape(London_Boundary)+
  tm_polygons("NAME",
              legend.show = FALSE,
              palette = 'black',
              alpha = 0.05)




# Map
ActuCap_Map <- tm_shape(ActuCap)+
  tm_bubbles("Total Flow",
             size = 0.8 ,
             scale = 0.5,
             style = "jenks",
             alpha = 0.7,
             col = "Total Flow",
             border.col = "black",
             title.col = "Actual Capicity",
             border.alpha = 0.3)




London_Base+ActuCap_Map
```

## Estimated Demand
### Distribution of Estimated Demand

```{r}
# Distribution
EstiCapDis <- ggplot(EstiFlow)+
  geom_histogram(aes(x = EstiFlow$PageRank),
                 fill = brewer.pal(7,"Blues")[7],
                 col =brewer.pal(7,"Greys")[2],
                 alpha = 0.8,
                 bins = 60)+
  scale_x_continuous()+
  scale_y_continuous()+
  labs(title="Distribution of Estimated Demand(PageRank(%))",
       x="PageRank(%)",
       y="Frequency")





EstiCapDis
```


### Transformed Distribution of Actual Demand

```{r}
# Transformed Distribution
EstiCapDis_Tran <- ggplot(EstiFlow)+
  geom_histogram(aes(x = log10(EstiFlow$PageRank)),
                 fill = brewer.pal(7,"Blues")[7],
                 col =brewer.pal(7,"Greys")[2],
                 alpha = 0.8,
                 bins = 60)+
  scale_x_continuous(breaks = seq(round(min(log10(EstiFlow$PageRank)),2),
                                  round(max(log10(EstiFlow$PageRank)),2),
                                  round((max(log10(EstiFlow$PageRank))-min(log10(EstiFlow$PageRank)))/10,1)))+
  scale_y_continuous()+
  labs(title="Distribution of log10((PageRank(%))",
       x="log10(PageRank(%))",
       y="Frequency")




EstiCapDis_Tran
```


### Estimated Demand on Map

```{r}
# Map
EstiCap_Map <- tm_shape(EstiCap)+
  tm_bubbles("PageRank",
             size = 0.8,
             scale = 0.5,
             style = "jenks",
             alpha = 0.7,
             col = "PageRank",
             title.col = "Estimated Capicity(PageRank)",
             breaks = c(0:6)*(max(EstiCap@data$PageRank)-min(EstiCap@data$PageRank))/6,
             border.col = "black",
             border.alpha = 0.3)




London_Base+EstiCap_Map
```

